---
title: "p8105_hw3_mc5503"
author: "mc5503"
date: '2023-10-08'
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
library(p8105.datasets)
data("instacart")
```

# Problem 1
The size of the dataset is `r nrow(instacart)` *`r ncol(instacart)`. Each order contains serveral products. One product belongs to one aisle and one department. For example, order 1 is in Thursday, the customer bought `r nrow(instacart |> filter(order_id==1))` things.`r instacart$product_name[1]` belongs to `r instacart$aisle[1]` and the `r instacart$department[1]` department.
```{r}
str(instacart)
```

## analysis of aisle
```{r}
aisle_n=instacart|>
  count(aisle,name="n_items")|>
  arrange(desc(n_items))
```
There are total `r nrow(aisle_n)` aisles and the `r aisle_n$aisle[1]` aisle is the most items ordered from. 

## number of items ordered in each aisle
```{r}
instacart |> 
  count(aisle,name="n_items")|>
  arrange(desc(n_items))|>
  filter(n_items>10000)|>
  mutate(aisle = forcats::fct_reorder(aisle, desc(n_items)))|>
  ggplot(aes(y=n_items,x=aisle))+
  geom_col()+
  theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = .5))
```

fresh fruits and vegetables are the two most popular aisles. 

## popular items
```{r}
bi=
  instacart|>
  filter(aisle=="baking ingredients")|>
  count(product_name, name = "n_items")|>
  arrange(desc(n_items))|>
  head(n=3L)
knitr::kable(bi)
dfc=
  instacart|>
  filter(aisle=="dog food care")|>
  count(product_name, name = "n_items")|>
  arrange(desc(n_items))|>
  head(n=3L)
knitr::kable(dfc)
pvf=
  instacart|>
  filter(aisle=="packaged vegetables fruits")|>
  count(product_name, name = "n_items")|>
  arrange(desc(n_items))|>
  head(n=3L)
knitr::kable(pvf)
```

The 3 most popular items in baking ingredients is `r bi$product_name`.The 3 most popular items in dog food care is `r dfc$product_name`. The 3 most popular items in packaged vegetables fruits is `r pvf$product_name`.

##  mean hour of the day
```{r}
instacart|>
  filter(product_name %in% c("Pink Lady Apples","Coffee Ice Cream"))|>
  group_by(product_name,order_dow)|>
  summarize(
    mean_hour_of_the_day=mean(order_hour_of_day))|>
  pivot_wider(
    names_from = order_dow,
    values_from = mean_hour_of_the_day
  )|> 
  rename("Sunday"="0","Monday"="1","Tuesday"="2","Wednesday"="3","Thursday"="4","Friday"="5","Saturdy"="6")|>
  knitr::kable(digits = 1)
```
Use `group_by` and `summarize` to calculate the mean hour of the day for each day in one week. In average, the mean hour of the day of Coffee Ice Cream is longer.

# Problem 2
```{r}
data("brfss_smart2010")
```

